---
title: "Course Project: Practical machine learning"
author: "RKT"
date: "12 February 2016"
output: html_document
---

### Setting up the work environment

Setting the  working directory
```{r}
setwd("~/coursera/practicalMachineLearning/")
```

Loading the required packages
```{r}
suppressWarnings(library(caret))
suppressWarnings(library(data.table))
```

Loading the data
```{r}
data <- read.table("pml-training.csv", header = TRUE, sep = ",")
quizData <- read.table("pml-testing.csv", header = TRUE, sep = ",")
```


### Data Cleaning

Understanding the data
```{r}
#names(data)
nrow(data)
table(data$classe)
#head(quizData)
#tail(quizData)
```

From the overview of the quizData, it appears that there are columns which have all the entries missing. Since, those columns can not be used for prediction, I have removed all those columns from both the data and the quizData. Also, there are columns containing timestamp or username (columns 1 to 5). Those data are also removed as they have no predictive power.

```{r}
index <- sapply(quizData, function(x) sum(is.na(x)) == nrow(quizData))
data <- data[,!(index)]
data <- data[, -c(1:5)]
quizData <- quizData[,!(index)]
quizData <- quizData[, -c(1:5)]
```

### Model training, evaluation and prediction

#### Model selection and training
I have chosen random forest algorithm because of its accuracy and efficiency. Since, the random forest algorithm has in-built algorithm for the estimation of unbiased classification error using out-of-bag (OOB) data, we therefore need not split the data into training and the test data set.

```{r, cache=TRUE}
rfModel <- train(classe ~ ., data = data, method = "rf")
```

#### Model evaluation
```{r}
print(rfModel)
```

We see that the accuracy (out-of-bag (oob))  is 0.9974 (mtry = 28). In the calculation of the oob accuracy, prediction of only those trees that did not use oob data for training are used. However, in the out-of-sample error calculations all the trees will be used. I, therefore, expect out-of-sample error to be smaller than 0.0026 (1-0.9974).

#### Model prediction
```{r}
predict(rfModel, quizData)
```

The model got all the predictions right for the quiz question. This provides proof that out-of-sample error is, indeed, smaller than 0.0027 


